<!-- front page -->

```{r include = FALSE}
source("code/common.R")
library(tidyverse)
library(devtools)
library(DT)
library(vembedr)
library(knitr)
library(readr)
library(gt)
# devtools::install_github("hadley/emo")

# parameters

class_days <- c("Tuesday and Thursday")
start_time <- "12 p.m. EST."
start_day  <- "Monday"
```

```{r links, child="admin/md/links.md"}
```

## Mason Notes {.unnumbered}

This website is constantly changing. I've included the source code for this website in the class [repo][course_repo]. I encourage you to contribute to the [course code][course_git]. If you catch typos or errors, please issue a pull request with the fixes. If you find cool or useful resources, please add them. By the end of the semester, I would love for everyone to have contributed to the course materials.

### How to use these notes {.unnumbered}

To navigate these course notes, use the table of contents on the left side of the screen. You can open or close the table of contents using the hamburger icon (horizontal bars) at the top of the document. Additionally, there are other icons at the top of the document for searching within the text, and for adjusting the size, font, or color scheme of the page. The document will be updated (unpredictably) throughout the semester.

Each module corresponds to a week's worth of material. Most modules are dedicated to improving a specific skill or at the very least dedicated to a specific theme. Within each module, you will find a range of resources including embedded videos, slides, activities, labs, and tutorials. The skills developed in each module are designed to build upon those you've learned in previous modules.  Eventually, this class will have more modules available than weeks in a semester, so that you -- the reader -- can choose your own adventure (err... module) you'd like to start with.

Although these notes have some textbook-like features, they are neither comprehensive nor completely original. The main purpose is to give you all a set of common materials to draw upon during the course. In class, we will sometimes do things outside the notes. The idea here is that you will read the materials and try to learn from them, just as you will attend classes and try to learn from them.


# Attribution {.unnumbered}

This class leans heavily on other peoples' materials and ideas. I have done my best to document the origin of the materials and ideas. In particular, I have noted people whose work has been a major contribution as well as those who have additional contributions. You can see specific changes by examining the [edit history on the git repo][edits]

## Major Attributions {.unnumbered}

* Mason Garrison's [Data Science for Psychologists][ds4p]
* Jenny Bryan's ([jennybryan]) [STAT 545][stat-545]
* Joe Rodgers's 2101
* Alisa Beyer [Introduction to Statistics for Psychology](https://open.maricopa.edu/psy230mm/)
* Danielle Navarro's [Learning Statistics with R](https://learningstatisticswithr.com/)

## Additional Attributions {.unnumbered}

* Eric Stone's 210
* Eranda Jayawickreme's 310

# License {#license .unnumbered}

This work is licensed under a [Creative Commons Attribution-ShareAlike 4.0 International License](https://creativecommons.org/licenses/by-sa/4.0/).

<center>
<i class="fab fa-creative-commons fa-2x"></i><i class="fab fa-creative-commons-by fa-2x"></i><i class="fab fa-creative-commons-sa fa-2x"></i>
</center>

*This information is a human-readable summary of (and not a substitute for) the license.
Please see <https://creativecommons.org/licenses/by-sa/4.0/legalcode> for the full legal text.*

**You are free to:**

- **Share**---copy and redistribute the material in any medium or
  format

- **Remix**---remix, transform, and build upon the material for any
  purpose, even commercially.

The licensor cannot revoke these freedoms as long as you follow the
license terms.

**Under the following terms:**

- **Attribution**---You must give appropriate credit, provide a link
  to the license, and indicate if changes were made. You may do so in
  any reasonable manner, but not in any way that suggests the licensor
  endorses you or your use.
  
- **ShareAlike**---If you remix, transform, or build upon the material, you must distribute your contributions under the same license as the original. 

- **No additional restrictions**---You may not apply legal terms or
  technological measures that legally restrict others from doing
  anything the license permits.

**Notices:**

You do not have to comply with the license for elements of the
material in the public domain or where your use is permitted by an
applicable exception or limitation.

No warranties are given. The license may not give you all of the
permissions necessary for your intended use. For example, other rights
such as publicity, privacy, or moral rights may limit how you use the
material.

```{r include = FALSE}
# automatically create a bib database for R packages
knitr::write_bib(
  c(
    .packages(),
    "bookdown",
    "knitr",
    "rmarkdown"
  ),
  "packages.bib"
)
```



# Colophon {-}

These notes was written in [bookdown](http://bookdown.org/) inside [RStudio](http://www.rstudio.com/ide/). The [website][web] is hosted with [github](https://www.github.com), The complete source is available from [github][git].

The book style is based on a design by [Desirée De Leon](https://desiree.rbind.io/).

This version of the notes was built with:


```{r message = FALSE,error=TRUE, warning = FALSE, echo = FALSE}
# needed because new_session is set to true in _bookdown.yml
all_pkgs <- renv::dependencies(path = "DESCRIPTION") %>%
  pull(Package)

session <- devtools::session_info(pkgs = all_pkgs)

session$platform
```

Along with these packages:

```{r message = FALSE,error=TRUE, warning = FALSE, echo = FALSE}
session$packages %>%
  select(package, loadedversion, date, source) %>%
  DT::datatable(
    rownames = FALSE,
    class = "cell-border stripe",
    filter = list(position = "top"),
    options = list(
      pageLength = nrow(session$packages),
      autoWidth = TRUE,
      bInfo = FALSE,
      paging = FALSE
    )
  )
```
