# Guidance {-}

```{r include = FALSE}
source("code/common.R")
```

```{r links, child="admin/md/links.md"}
```

<!-- To Do: find a way to embed a pdf in rmarkdown -->

## Syllabus

This document is not the syllabus. Please refer to Mason Garrison's syllabus here: https://smasongarrison.github.io/syllabi/ 


## Materials

### Hardware

This class requires that you have a laptop that can run R.

### Required Texts

The text is intended to supplement the videos, lecture notes, tutorials, activities, and labs. You probably need to consume all of them in order to be successful in this class.



### Software

#### R and RStudio

```{r, echo=FALSE}
embed_url("https://www.youtube.com/watch?v=kVIZGCT5p9U") %>%
 use_align("center")
```

R is a free software environment for statistical computing and graphics. It compiles and runs on a wide variety of UNIX platforms, Windows, and MacOS.

RStudio is a free _integrated development environment_ (IDE), a powerful user interface for R.

#### Git and Github

Git is a version control system. Its original purpose was to help groups of developers work collaboratively on big software projects. Git manages the evolution of a set of files – called a repository – in a structured way. Think of it like the "Track Changes" features from Microsoft Word.

Github is a free IDE and hosting service for Git. As a Wake Forest student, you should be able to access the [GitHub Student Developer Pack for free](https://education.github.com/benefits?type=student). It includes a free PRO upgrade for your github account.


